---
title: "Summary simulation results"
author: ""
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    number_sections: true
    code_folding: hide 
    theme: readable
    # journal, readable, paper 
---

```{r setup, include=FALSE}
library(ggplot2)
library(gridExtra)
library(ggpubr)
library(tidyverse) 

# Datasets
load("C:/Users/mbofi/Dropbox/C5/Scripts/GitKraken/CBE_selection/results/results_bias.RData")
dataset_ARE= dataset
load("C:/Users/mbofi/Dropbox/C5/Scripts/GitKraken/CBE_selection/results/results_SS.RData")
dataset_SS= dataset
load("C:/Users/mbofi/Dropbox/C5/Scripts/GitKraken/CBE_selection/results/results_SS_H0True.RData")
dataset_SSh0= dataset

dataset = data.frame(dataset_ARE,dataset_SS[,21:27],dataset_SSh0[,24:25])

knitr::opts_chunk$set(echo = TRUE) 
```

# Changelog

Main modifications with respect to previous versions:

2020-11-06

  - We compare two decision criteria: ARE and Ratio Sample Sizes.
  - We evaluate the power when one of the composite components has no treatment effect. See section 2.1.

2020-10-30

  - We have added new constraints on the possible correlation values. See section 2.3. 
  - Correction on the correlation estimator in the blinded approach. See overleaf document for further details.
  - Study on the bias of the correlation estimator. See section 3.3. 
  

2020-09-24

  - The scenarios we considered for the parameters ($p_1^{(0)},p_2^{(0)},OR_1,OR_2$) are a subset of the previous version, while for the correlation  ($\rho$) we have added more values. 
  - For the endpoint selection, we estimate the correlation at the end of the study based on two different approaches: i) using blinded data; and ii) using unblinded data. See section 2.2. for further explanation. We remark that in contrast with the previous version, we do not use the expected effect on the composite endpoint to estimate the correlation. 

# Design and main assumptions

We simulate a two-arm trial with two binary endpoints according to the following parameters:

  - Probability endpoint 1 control group ($p_1^{(0)}$): $0.1, 0.2$;
  - Probability endpoint 2 control group ($p_2^{(0)}$): $0.1, 0.25$;
  - Odds ratio endpoint 1 ($OR_1$): $0.6, 0.7, 0.8$;
  - Odds ratio endpoint 2 ($OR_2$): $0.75, 0.8$;
  - Correlation between endpoints ($\rho$): $0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8$.  

The correlation between the endpoints is assumed to be equal for both groups. We note that since the correlation is bounded between two values defined by ($p_1^{(0)},p_2^{(0)},OR_1,OR_2$), not all the combinations of  ($p_1^{(0)},p_2^{(0)},OR_1,OR_2,\rho$)  are possible. We calculate the correlation bounds for each combination of ($p_1^{(0)},p_2^{(0)},OR_1,OR_2$) and discard those scenarios in which the correlation is not within the valid range. The total number of scenarios is $183$.

We consider the composite binary endpoint formed by endpoints 1 and 2. 
Based on the parameters ($p_1^{(0)},p_2^{(0)},OR_1,OR_2,\rho$), we compute:

  - Probability composite endpoint control group ($p_*^{(0)}$);
  - Odds ratio composite endpoint ($OR_*$);
  
The sample size used for the simulation is calculated to have $0.80$ power to detect an effect of $OR_1$ on the endpoint 1 at significance level $\alpha = 0.05$. 

## Decision criterias  
  
We assume that the endpoint 1 is more relevant for the clinical question than endpoint 2.


**ARE Method**:

We consider the Asymptotic Relative Efficiency (ARE) method to quantify the efficiency of using the composite endpoint over the endpoint 1 as primary endpoint of the trial. Based on the parameters ($p_1^{(0)},p_2^{(0)},OR_1,OR_2,\rho$), we compute:

  - ARE value;
  - ARE Decision: *CE*, if the composite endpoint is preferred; *RE*, if the endpoint 1 is preferred.

**Ratio Sample sizes**:

We calculate the sample size to have $0.80$ power to detect an effect of $OR_1$ on the endpoint 1, given the probability $p_1^{(0)}$ under the control group,  at significance level $\alpha = 0.05$, say $n_1$. 

We also calculate the sample size  to have $0.80$ power to detect an effect of $OR_*$ on the composite endpoint, given the probability $p_*^{(0)}$ under the control group, at significance level $\alpha = 0.05$, say $n_*$. Note that for calculating the sample size for the composite endpoint have been used the values of the parameters ($p_1^{(0)},p_2^{(0)},OR_1,OR_2,\rho$).

The decision criteria is the ratio of the two sample sizes:

$$n_1 / n_*$$

so that if the ratio is smaller than $1$, we use the relevant endpoint; whereas if the ratio is larger than $1$, we use the composite endpoint.


## Cases under study

We evaluate through simulations the statistical power and significance level under different situations and using different primary endpoints.

For each scenario, we compare the two groups based on three endpoints:

  - the relevant endpoint: we consider endpoint 1 as the primary endpoint;
  - the composite endpoint: we consider the composite endpoint formed by the union of endpoint 1 and endpoint 2 as the primary endpoint;
  - the selected endpoint: we select the primary endpoint at the end of the trial. We do so based on the ARE method and using blinded/unblinded data to estimate the probabilities under the control group and the correlation. 
  
Remark:   we simulate $100000$ trials for each scenario and for each endpoint considered as primary endpoint; 
<!-- 2) we estimate the correlation using the estimated probabilities ($p_1^{(0)},p_2^{(0)},p_*^{(0)}$).  -->

For each scenario, we evaluate the following cases:

  1. Under alternative hypothesis ($H_1$):
  
      a. testing the group differences based on the endpoint 1.
      b. testing the group differences based on the composite endpoint. 
      c. testing the group differences based on the selected endpoint.
  
  2. Under the null hypothesis ($H_0$): 
  
      a. testing the group differences based on the endpoint 1.
      b. testing the group differences based on the composite endpoint. 
      c. testing the group differences based on the selected endpoint.

When simulating under $H_0$, we consider $OR_1=OR_2=1$; whereas when simulating under $H_1$,
we consider the following situations:
<!-- ; we consider the previous mentioned values for $OR_1$, $OR_2$, and $OR_*$. -->

  - Effect on both composite components: That is, we consider the previous mentioned values for $OR_1$, $OR_2$, and the resulting $OR_*$.
  - Effect on only one component: That is, we plan the trial based on the previous mentioned values $OR_1$, $OR_2$, and the resulting $OR_*$, but we simulate that $OR_1=1$ or $OR_2=1$.

## Estimation correlation and ARE recalculation

We consider two different approaches for estimating the correlation between the endpoints 1 and 2.

**Blinded approach**:

Based on the observed responses in the pooled sample, we estimate the probabilities $p_1$, $p_2$, and $p_*$, where $p_k=\pi p_k^{(0)} + (1-\pi)p_k^{(1)}$ for $k=1,2,*$  and $\pi=n^{(0)}/n$.
Assuming the expected effects for the endpoints 1 and 2  (say $OR_1$ and $OR_2$) pre-specified in advance, we obtain estimates of the probabilities of each composite component under the control group $p_1^{(0)},p_2^{(0)}$ and subsequently the estimates of the probabilities under the  treatment group $p_1^{(1)},p_2^{(1)}$.

Using the estimated probabilities for each composite component in each group ($p_1^{(0)},p_2^{(0)},p_1^{(1)},p_2^{(1)}$) and using the estimated pooled probability of the composite endpoint ($p_*$), we get an estimate of the correlation.

We calculate the ARE by using the estimated probabilities $p_1^{(0)},p_2^{(0)}$ and the estimated correlation.


**Unblinded approach**:

Based on the observed responses in patients in the control group, we estimate the probabilities $p_1^{(0)},p_2^{(0)},p_*^{(0)}$. Then, using these estimated probabilities, we obtain an estimate of the correlation between the Endpoints 1 and 2 in the control group $\rho^{(0)}$.

We do the same for the treatment group, and then we obtain an estimate of the correlation in the treatment group $\rho^{(1)}$.

We calculate the ARE by using the estimated probabilities $p_1^{(0)},p_2^{(0)}$ and the mean of the two estimates of the correlation.


## Correlation bounds

As mentioned before, 
Pearson correlation coefficient between two binary endpoints takes values between two bounds: denote by $B_L(\cdot)$ the lower bound, and by $B_U(\cdot)$ the upper bound. Both $B_L(\cdot)$ and $B_U(\cdot)$ are functions of  the marginal probabilities of the binary endpoints.

Thus, in our context, the correlation between endpoints 1 and 2 in the control group lies within $B_L(p_1^{(0)},p_2^{(0)})$ and $B_U(p_1^{(0)},p_2^{(0)})$; whereas the correlation between them in the treatment group takes values between  $B_L(p_1^{(1)},p_2^{(1)})$ and $B_U(p_1^{(1)},p_2^{(1)})$.

Since we assume that the correlation is the same in both treatment groups, we require that:

$$
max(B_L(p_1^{(0)},p_2^{(0)}),B_L(p_1^{(1)},p_2^{(1)})) \leq \rho \leq min(B_U(p_1^{(0)},p_2^{(0)}),B_U(p_1^{(1)},p_2^{(1)}))
$$

In the estimation approaches explained above, we compute the correlation bounds based on the estimated probabilities and force the estimated correlation to take values within this interval. 
We notice that this interval can be different from the one calculated based on the true probabilities (those used to run the simulations).
<!-- , and therefore be narrower or wider depending on the circumstances. -->


<!-- Note: Future steps, consider different association measures between endpoints E1 and E2, say for example Relative Overlap or Index Jaccard.  -->

# Results

## Under alternative hypothesis ($H_1$)

### Cases (1a) and (1b)

In what follows we compare the statistical powers obtained using the composite endpoint with the ones obtained using the endpoint 1 as  primary endpoint. We also evaluate the concordance between the choice of the primary endpoint
<!-- decision between using the composite endpoint or the endpoint 1  -->
-according to the ARE value and the ratio of the sample sizes- and the difference in powers obtained in the simulation.  In this case, we calculate the ARE at the beginning of the study, and then according to the parameters ($p_1^{(0)},p_2^{(0)},OR_1,OR_2,\rho$).

```{r} 
dataset$diff_powers = dataset$Test_Power_CE - dataset$Test_Power_RE 
```

The following plot shows the relationship between the ARE value and the difference in power between using the composite endpoint (CE) or using the endpoint 1 (RE). 
```{r}
ggplot(dataset, aes(x=ARE, y=diff_powers, color=as.factor(decision))) + geom_point(size=2) + labs(y = "Difference in power (Power CE - Power RE)", x="ARE", color="Decision")
```

Cases in which there is no concordance between the decision and the power gain: 
```{r}
dataset$gain_power = ifelse(dataset$diff_powers>0, "CE", "RE")
dataset$gain_power = as.factor(dataset$gain_power)
dataset$decision = as.factor(dataset$decision)

summary(subset(dataset,dataset$gain_power != dataset$decision)[,c(3,4,18,19,20,21,22)])
```

### Case (1c) 

The following figure illustrates the empirical power we obtained using the composite endpoint (CE), the endpoint 1 (RE) or the selected endpoint (SE). 
```{r, warning=F}

subd=dataset#subset(dataset,is.na(dataset$Test_Power_ES_ub)==F)

power_data <- data.frame(Power=c(dataset$Test_Power_ES,dataset$Test_Power_ES_ub,dataset$Test_Power_CE,dataset$Test_Power_RE),
                         Endpoint=c(rep("SE",length(dataset$Test_Power_ES)),
                                    rep("SE_ub",length(dataset$Test_Power_ES_ub)),
                                    rep("CE",length(dataset$Test_Power_CE)),
                                    rep("RE",length(dataset$Test_Power_RE)))
)

ggplot(power_data, aes(x=Endpoint, y=Power)) + geom_boxplot()
```

<!-- We compare the maximum between the power using the composite endpoint and the maximum using the relevant endpoint with the power obtained using the selected endpoint. -->

<!-- ```{r warning=FALSE} -->

<!-- dataset$diff_es = ifelse(dataset$gain_power=="CE", dataset$Test_Power_ES-dataset$Test_Power_CE,dataset$Test_Power_ES-dataset$Test_Power_RE) -->
<!-- summary(dataset$diff_es) -->
<!-- ``` -->

### Comparing empirical powers

For each combination of ($p_1^{(0)},p_2^{(0)},OR_1,OR_2$), we plot the empirical power for each design -Selected Endpoint with blinded estimation (SE), Selected Endpoint with unblinded estimation (SE_u), Composite Endpoint (CE), and Relevant Endpoint (RE)- according to the correlation $\rho$. 
In the empirical powers for the Selected Endpoint, we plot the points in different colors depending on which  endpoint would have been chosen using  the ARE calculated from the *true* parameters ($p_1^{(0)},p_2^{(0)},OR_1,OR_2,\rho$).

For each combination of ($p_1^{(0)},p_2^{(0)},OR_1,OR_2$),
we also present a table with: 

  1. **Corr**: The values considered for the correlation in this scenario;
  2. **ARE**: The value of the ARE computed using the parameters used in the design, i.e., ($p_1^{(0)},p_2^{(0)},OR_1,OR_2$) and $\rho$;
  3. **%CE**: The percentage of cases over the 100000 replications in which the selected endpoint was the composite endpoint using blinded data; 
  4. **%CE_u**: The percentage of cases over the 100000 replications in which the selected endpoint was the composite endpoint using unblinded data.

```{r, fig.height=4, fig.width=10, warning=FALSE} 
p <- list() 
enum = 1 
it <- 1
for(i in 1:max(dataset$scenario)){
  sub=subset(dataset,dataset$scenario==i)
  # 
  p[[enum]] <-ggplot(sub, aes(x=corr, y=Test_Power_ES, color=as.factor(decision)))  +
    geom_point(size=2)+ ggtitle(paste("Scenario", dataset$scenario[it], "\n (p1,p2,OR1,OR2) \n=(", dataset$p0_e1[it],",",dataset$p0_e2[it],",",dataset$OR1[it],",",dataset$OR2[it],")"))+geom_point(size=2)  + labs(y = "Empirical Power (ES)", x="Correlation", color="Decision") + coord_cartesian(ylim = c(0.60, 1))+ geom_path()+ theme(plot.title = element_text(size=9),legend.position = c(0.8, 0.2))
  p[[enum+1]] <-ggplot(sub, aes(x=corr, y=Test_Power_ES_ub, color=as.factor(decision)))  +
    geom_point(size=2)+ ggtitle(paste("Scenario", dataset$scenario[it], "\n (p1,p2,OR1,OR2) \n=(", dataset$p0_e1[it],",",dataset$p0_e2[it],",",dataset$OR1[it],",",dataset$OR2[it],")"))+geom_point(size=2)  + labs(y = "Empirical Power (ES, unblinded)", x="Correlation", color="Decision") + coord_cartesian(ylim = c(0.60, 1))+ geom_path()+ theme(plot.title = element_text(size=9),legend.position = c(0.8, 0.2))
  # 
  p[[enum+2]] <-ggplot(sub, aes(x=corr, y=Test_Power_ES_bSS, color=as.factor(ss_decision)))  +
    geom_point(size=2)+ ggtitle(paste("Scenario", dataset$scenario[it], "\n (p1,p2,OR1,OR2) \n=(", dataset$p0_e1[it],",",dataset$p0_e2[it],",",dataset$OR1[it],",",dataset$OR2[it],")"))+geom_point(size=2)  + labs(y = "Empirical Power (ES)", x="Correlation", color="Decision SS") + coord_cartesian(ylim = c(0.60, 1))+ geom_path()+ theme(plot.title = element_text(size=9),legend.position = c(0.8, 0.2))
  p[[enum+3]] <-ggplot(sub, aes(x=corr, y=Test_Power_ES_ubSS, color=as.factor(ss_decision)))  +
    geom_point(size=2)+ ggtitle(paste("Scenario", dataset$scenario[it], "\n (p1,p2,OR1,OR2) \n=(", dataset$p0_e1[it],",",dataset$p0_e2[it],",",dataset$OR1[it],",",dataset$OR2[it],")"))+geom_point(size=2)  + labs(y = "Empirical Power (ES, unblinded)", x="Correlation", color="Decision SS") + coord_cartesian(ylim = c(0.60, 1))+ geom_path()+ theme(plot.title = element_text(size=9),legend.position = c(0.8, 0.2))
  # 
  p[[enum+4]] <-ggplot(sub, aes(x=corr, y=Test_Power_CE))  +
    geom_point(size=2)+ ggtitle(paste("Scenario", dataset$scenario[it], "\n (p1,p2,OR1,OR2) \n=(", dataset$p0_e1[it],",",dataset$p0_e2[it],",",dataset$OR1[it],",",dataset$OR2[it],")"))+geom_point(size=2)  + labs(y = "Empirical Power (CE)", x="Correlation", color="Decision") + coord_cartesian(ylim = c(0.60, 1))+ geom_path()+ theme(plot.title = element_text(size=9))
  p[[enum+5]] <-ggplot(sub, aes(x=corr, y=Test_Power_RE))  +
    geom_point(size=2)+ ggtitle(paste("Scenario", dataset$scenario[it], "\n (p1,p2,OR1,OR2) \n=(", dataset$p0_e1[it],",",dataset$p0_e2[it],",",dataset$OR1[it],",",dataset$OR2[it],")"))+geom_point(size=2)  + labs(y = "Empirical Power (RE)", x="Correlation", color="Decision") + coord_cartesian(ylim = c(0.60, 1))+ geom_path()+ theme(plot.title = element_text(size=9))

  stable <- data.frame(Corr=sub$corr, 
                       ARE=round(sub$ARE,2),"%CE"=round(100*sub$decision_ES,2),
                       "%CE_u"=round(100*sub$decision_ES_ub,2),check.names=FALSE)  

  p[[enum+6]] <- ggtexttable(stable, rows = NULL, theme = ttheme(base_style = "default", base_size = 9))
  
  stable <- data.frame(Corr=sub$corr, 
                       SS=round(sub$ss_ratio,2),"%CE"=round(100*sub$decision_ES_bSS,2),
                       "%CE_u"=round(100*sub$decision_ES_ubSS,2),check.names=FALSE)  

  p[[enum+7]] <- ggtexttable(stable, rows = NULL, theme = ttheme(base_style = "default", base_size = 9))

  enum=enum+8 
  it <- it + dim(subset(dataset,dataset$scenario==i))
}
marrangeGrob(p,ncol=4,nrow=1,top=NULL) 
# ,heights = c(5, 10)
```

<!-- ### Toy example to compare blinded and unblinded approaches -->

## Under the null hypothesis ($H_0$)

### Cases (2a), (2b), and (2c) 


The following boxplots show the significance level obtained using the composite endpoint (CE), the endpoint 1 (RE) or the selected endpoint (SE). 
```{r, warning=F}
signlevel_data <- data.frame(signlevel=c(dataset$Test_Reject_ES,
                                         dataset$Test_Reject_ES_ub,
                                         dataset$Test_Reject_CE,dataset$Test_Reject_RE),
                             Endpoint=c(rep("SE",length(dataset$Test_Power_ES)),
                                        rep("SE(ub)",length(dataset$Test_Reject_ES_ub)),
                                        rep("CE",length(dataset$Test_Power_CE)),
                                        rep("RE",length(dataset$Test_Power_RE)))
                             )

ggplot(signlevel_data, aes(x=Endpoint, y=signlevel)) + geom_boxplot()
```


```{r}
summary(dataset[,27:30])
```

Summary cases type 1 error rate$>0.055$ with SE(ub):
```{r}
# summary(subset(dataset,dataset[,30]>0.055)) 
subset(dataset,dataset[,30]>0.055)[,c(1,2,3,4,14,15,16,17,18)]
```

## On the bias of the correlation estimator

We ran an extra simulation to assess the correlation estimator in both blinded and unblinded approaches. Using the same scenarios we considered before, we simulate each scenario $nsim = 1000$ times. For each run, we estimate the correlation using the blinded and unblinded approach. We then compute the mean and variance of these estimators. We compute the bias as the difference between the mean value and the true value of the correlation in that scenario. 

Scatterplots estimated correlation vs true correlation:
```{r}
ggplot(dataset, aes(x=corr, y=corr_est_ub_mean)) + geom_point(size=2) + labs(y = "Mean value correlation estimates (Unblinded approach)", x="True correlation")

ggplot(dataset, aes(x=corr, y=corr_est_b_mean)) + geom_point(size=2) + labs(y = "Mean value correlation estimates (Blinded approach)", x="True correlation") 
```

Summary bias:
```{r} 
# summary(dataset$corr)
# summary(dataset[,31:36])
summary(dataset[,35:36])
```

# Decision criteria

```{r}
summary(dataset)
```


```{r}
dataset$diff_powers = dataset$Test_Power_CE - dataset$Test_Power_RE 
ggplot(dataset, aes(x=ARE, y=diff_powers, color=as.factor(decision))) + geom_point(size=2) + labs(y = "Difference in power (Power CE - Power RE)", x="ARE", color="Decision")
```


```{r}
# dataset$diff_powers = dataset$Test_Power_CE - dataset$Test_Power_RE 
ggplot(dataset, aes(x=ss_ratio, y=diff_powers, color=as.factor(ss_decision))) + geom_point(size=2) + labs(y = "Difference in power (Power CE - Power RE)", x="Ratio SS", color="Decision")
```

# Conclusions and future directions

Conclusions:

  - The powers using the selected endpoint (both blinded and unblinded approaches) are in general larger than the ones obtained using the composite endpoint and the relevant endpoint.  
  - The significance level is similar in the four situations.

<!-- Open questions: -->

<!--   - Review assumptions correlation bounds.  -->
<!--   - Consider different association measures between endpoints E1 and E2, say for example Relative Overlap or  Jaccard Index.  -->
<!--   - Improve R code -->

Future steps: 

  - Consider an interim stage to estimate the probabilities under the control group and the correlation.



